---
title: "CellChat"
author: "GinoBonazza (ginoandrea.bonazza@usz.ch)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r knitr config, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(warning = FALSE)

knitr::opts_chunk$set(message = FALSE)

knitr::opts_chunk$set(cache = FALSE)

knitr::opts_chunk$set(dpi = 300, fig.align = "center")
```

## Setup

```{r setup, class.source = "fold-hide"}
# Get current file name to make folder
current_file <- "CellChat"

# Load libraries
library(here)
library(readr)
library(readxl)
library(xlsx)
library(Seurat)
library(DropletUtils)
library(Matrix)
library(scDblFinder)
library(scCustomize)
library(dplyr)
library(ggplot2)
library(magrittr)
library(tidyverse)
library(reshape2)
library(S4Vectors)
library(SingleCellExperiment)
library(pheatmap)
library(png)
library(gridExtra)
library(knitr)
library(scales)
library(RColorBrewer)
library(Matrix.utils)
library(tibble)
library(ggplot2)
library(scater)
library(patchwork)
library(statmod)
library(ArchR)
library(clustree)
library(speckle)
library(ggplot2)
library(scater)
library(gprofiler2)
library(clusterProfiler)
library(org.Hs.eg.db)
library(AnnotationHub)
library(ReactomePA)
library(speckle)
library(reactable)
library(CellChat)
library(patchwork)

#Output paths
output_dir_data <- here::here("output", current_file)
if (!dir.exists(output_dir_data)) dir.create(output_dir_data)

if (!dir.exists(here::here("docs", "figure"))) dir.create(here::here("docs", "figure"))

output_dir_figs <- here::here("docs", "figure", paste0(current_file, ".Rmd"))
if (!dir.exists(output_dir_figs)) dir.create(output_dir_figs)
```


```{r}
iPS_integrated <- readRDS(here::here("output","QC_integration_annotation_Ctr", "iPS_integrated.rds"))
```

```{r options}
options(stringsAsFactors = FALSE)
```

```{r data.input}
data.input <- iPS_integrated[["RNA"]]@data # normalized data matrix
# For Seurat version >= “5.0.0”, get the normalized data via `seurat_object[["RNA"]]$data`
labels <- Idents(iPS_integrated)
meta <- data.frame(labels = labels, row.names = names(labels)) # create a dataframe of the cell labels
```

```{r createCellChat}
cellchat <- createCellChat(object = iPS_integrated, group.by = "ident", assay = "RNA")
cellchat <- addMeta(cellchat, meta = meta)
cellchat <- setIdent(cellchat, ident.use = "labels") # set "labels" as default cell identity
levels(cellchat@idents) # show factor levels of the cell labels
groupSize <- as.numeric(table(cellchat@idents)) # number of cells in each cell group
groupSize
```

```{r CellChatDB}
CellChatDB <- CellChatDB.human 
cellchat@DB <- CellChatDB
```

## Whole CellChat database

```{r}
# subset the expression data of signaling genes for saving computation cost
cellchat <- subsetData(cellchat) # This step is necessary even if using the whole database
#future::plan("multisession", workers = 4) # do parallel
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
#> The number of highly variable ligand-receptor pairs used for signaling inference is 692
```

```{r}
cellchat <- computeCommunProb(cellchat, type = "triMean")
cellchat <- filterCommunication(cellchat, min.cells = 10)
```

```{r}
df.net <- subsetCommunication(cellchat, slot.name = "net")
```


```{r}
cellchat <- computeCommunProbPathway(cellchat)
```

```{r}
cellchat <- aggregateNet(cellchat)
```

```{r}
df.pathways <- subsetCommunication(cellchat, slot.name = "netP", thresh = 0.05)
```

### Aggregated cell-cell communication network

```{r, fig.width=9, fig.height=3}
groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1,2), mar = c(0, 0, 2, 0), xpd=TRUE)
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")
```

### Cell-cell communication network split by sender
```{r, fig.width=9, fig.height=5}
mat <- cellchat@net$weight
par(mfrow = c(2,3), mar = c(0, 4, 0, 4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
```


### Interaction strength and number of interactions for each cell type pair

```{r}
netVisual_heatmap(cellchat, color.heatmap = "Reds", slot.name = "net", measure = "weight")
```

```{r}
netVisual_heatmap(cellchat, color.heatmap = "Reds", slot.name = "net", measure = "count")
```


### Dominant senders (sources) and receivers (targets)

```{r}
# Run the analysis without parallel processing
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP")

gg1 <- netAnalysis_signalingRole_scatter(cellchat)
gg1
```


### Signals contributing the most to outgoing or incoming signaling

```{r, fig.height=13, fig.width=11}
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", width = 10,
  height = 25)
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", width = 10,
  height = 25)
ht1 + ht2
```

```{r, fig.height=6, fig.width=8}
df <- rankNet(cellchat, mode = "single", stacked = T, return.data = T)$signaling.contribution %>%
  arrange(desc(contribution))
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", width = 7,
  height = 8, signaling = df[1:25,]$name)
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", width = 7,
  height = 8, signaling = df[1:25,]$name)
ht1 + ht2
```


### Top interactions between all cell types

```{r, fig.height=6, fig.width=6}
net <- df.net %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 1, legend.pos.y = 130, legend.pos.x = 5, slot.name = "net", net = net[1:25,])
```

```{r, fig.height=6, fig.width=6}
net <- df.pathways %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 0.65, legend.pos.y = 110, legend.pos.x = 5, slot.name = "netP", net = net[1:25,], small.gap = 2)
```


### Top interactions with endothelial cells as senders

```{r, fig.height=6, fig.width=6}
net <- df.net %>%
  filter(source == "Endothelial cells") %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 1, legend.pos.y = 130, legend.pos.x = 5, slot.name = "net", net = net[1:25,], sources.use = "Endothelial cells")
```

```{r, fig.height=8, fig.width=10}
net <- df.pathways %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, sources.use = "Endothelial cells", lab.cex = 1,legend.pos.y = 150, slot.name = "netP", net = net[1:100,])
```

```{r}
net <- df.net %>%
  filter(source == "Endothelial cells") %>%
  arrange(desc(prob))
df <- as.data.frame(net[1:25,]$interaction_name)
colnames(df) <- "interaction_name"
netVisual_bubble(cellchat, remove.isolate = FALSE, pairLR.use = df, sources.use = "Endothelial cells", angle.x = 45)
```


### Top interactions with endothelial cells as senders and cardiomyocytes as receivers

```{r, fig.height=6, fig.width=6}
net <- df.net %>%
  filter(source == "Endothelial cells") %>%
  filter(target %in% c("Ventricular Cardiomyocytes", "Atrial Cardiomyocytes")) %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 1, legend.pos.y = 135, legend.pos.x = 5, slot.name = "net", net = net[1:25,], sources.use = "Endothelial cells", targets.use = c("Ventricular Cardiomyocytes", "Atrial Cardiomyocytes"))
```

```{r}
df <- as.data.frame(net[1:25,]$interaction_name)
colnames(df) <- "interaction_name"
netVisual_bubble(cellchat, remove.isolate = FALSE, pairLR.use = df, sources.use = "Endothelial cells", angle.x = 45, targets.use = c("Ventricular Cardiomyocytes", "Atrial Cardiomyocytes"))
```

```{r}
saveRDS(cellchat, 
        here::here(output_dir_data, "iPS_cellchat.rds"))
```


## Subset only Secreted signaling

```{r}
cellchat <- createCellChat(object = iPS_integrated, group.by = "ident", assay = "RNA")
cellchat <- addMeta(cellchat, meta = meta)
cellchat <- setIdent(cellchat, ident.use = "labels") # set "labels" as default cell identity
levels(cellchat@idents) # show factor levels of the cell labels
groupSize <- as.numeric(table(cellchat@idents)) # number of cells in each cell group
groupSize
```

```{r}

CellChatDB <- CellChatDB.human 
cellchat@DB <- CellChatDB
CellChatDB.use <- subsetDB(CellChatDB, search = "Secreted Signaling", key = "annotation") # use Secreted Signaling
cellchat@DB <- CellChatDB.use
```

```{r}
# subset the expression data of signaling genes for saving computation cost
cellchat <- subsetData(cellchat, ) # This step is necessary even if using the whole database
#future::plan("multisession", workers = 4) # do parallel
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
#> The number of highly variable ligand-receptor pairs used for signaling inference is 692
```

```{r}
cellchat <- computeCommunProb(cellchat, type = "triMean")
cellchat <- filterCommunication(cellchat, min.cells = 10)
```

```{r}
df.net <- subsetCommunication(cellchat, slot.name = "net")
```


```{r}
cellchat <- computeCommunProbPathway(cellchat)
```

```{r}
cellchat <- aggregateNet(cellchat)
```

```{r}
df.pathways <- subsetCommunication(cellchat, slot.name = "netP", thresh = 0.05)
```

### Aggregated cell-cell communication network

```{r, fig.width=9, fig.height=3}
groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1,2), mar = c(0, 0, 2, 0), xpd=TRUE)
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")
```

### Cell-cell communication network split by sender
```{r, fig.width=9, fig.height=5}
mat <- cellchat@net$weight
par(mfrow = c(2,3), mar = c(0, 4, 0, 4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
```


### Interaction strength and number of interactions for each cell type pair

```{r}
netVisual_heatmap(cellchat, color.heatmap = "Reds", slot.name = "net", measure = "weight")
```

```{r}
netVisual_heatmap(cellchat, color.heatmap = "Reds", slot.name = "net", measure = "count")
```


### Dominant senders (sources) and receivers (targets)

```{r}
# Run the analysis without parallel processing
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP")

gg1 <- netAnalysis_signalingRole_scatter(cellchat)
gg1
```


### Signals contributing the most to outgoing or incoming signaling

```{r, fig.height=13, fig.width=11}
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", width = 10,
  height = 25)
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", width = 10,
  height = 25)
ht1 + ht2
```

```{r, fig.height=6, fig.width=8}
df <- rankNet(cellchat, mode = "single", stacked = T, return.data = T)$signaling.contribution %>%
  arrange(desc(contribution))
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", width = 7,
  height = 8, signaling = df[1:25,]$name)
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", width = 7,
  height = 8, signaling = df[1:25,]$name)
ht1 + ht2
```


### Top interactions between all cell types

```{r, fig.height=6, fig.width=6}
net <- df.net %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 1, legend.pos.y = 130, legend.pos.x = 5, slot.name = "net", net = net[1:25,])
```

```{r, fig.height=6, fig.width=6}
net <- df.pathways %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 0.65, legend.pos.y = 110, legend.pos.x = 5, slot.name = "netP", net = net[1:25,], small.gap = 2)
```


### Top interactions with endothelial cells as senders

```{r, fig.height=6, fig.width=6}
net <- df.net %>%
  filter(source == "Endothelial cells") %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 1, legend.pos.y = 130, legend.pos.x = 5, slot.name = "net", net = net[1:25,], sources.use = "Endothelial cells")
```

```{r, fig.height=8, fig.width=10}
net <- df.pathways %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, sources.use = "Endothelial cells", lab.cex = 1,legend.pos.y = 150, slot.name = "netP", net = net[1:100,])
```

```{r}
net <- df.net %>%
  filter(source == "Endothelial cells") %>%
  arrange(desc(prob))
df <- as.data.frame(net[1:25,]$interaction_name)
colnames(df) <- "interaction_name"
netVisual_bubble(cellchat, remove.isolate = FALSE, pairLR.use = df, sources.use = "Endothelial cells", angle.x = 45)
```


### Top interactions with endothelial cells as senders and cardiomyocytes as receivers

```{r, fig.height=6, fig.width=6}
net <- df.net %>%
  filter(source == "Endothelial cells") %>%
  filter(target %in% c("Ventricular Cardiomyocytes", "Atrial Cardiomyocytes")) %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 1, legend.pos.y = 135, legend.pos.x = 5, slot.name = "net", net = net[1:25,], sources.use = "Endothelial cells", targets.use = c("Ventricular Cardiomyocytes", "Atrial Cardiomyocytes"))
```

```{r}
df <- as.data.frame(net[1:25,]$interaction_name)
colnames(df) <- "interaction_name"
netVisual_bubble(cellchat, remove.isolate = FALSE, pairLR.use = df, sources.use = "Endothelial cells", angle.x = 45, targets.use = c("Ventricular Cardiomyocytes", "Atrial Cardiomyocytes"))
```

## Subset only Cell-Cell Contact signals

```{r}
cellchat <- createCellChat(object = iPS_integrated, group.by = "ident", assay = "RNA")
cellchat <- addMeta(cellchat, meta = meta)
cellchat <- setIdent(cellchat, ident.use = "labels") # set "labels" as default cell identity
levels(cellchat@idents) # show factor levels of the cell labels
groupSize <- as.numeric(table(cellchat@idents)) # number of cells in each cell group
groupSize
```

```{r}

CellChatDB <- CellChatDB.human 
cellchat@DB <- CellChatDB
CellChatDB.use <- subsetDB(CellChatDB, search = "Cell-Cell Contact", key = "annotation") # use Secreted Signaling
cellchat@DB <- CellChatDB.use
```

```{r}
# subset the expression data of signaling genes for saving computation cost
cellchat <- subsetData(cellchat, ) # This step is necessary even if using the whole database
#future::plan("multisession", workers = 4) # do parallel
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
#> The number of highly variable ligand-receptor pairs used for signaling inference is 692
```

```{r}
cellchat <- computeCommunProb(cellchat, type = "triMean")
cellchat <- filterCommunication(cellchat, min.cells = 10)
```

```{r}
df.net <- subsetCommunication(cellchat, slot.name = "net")
```


```{r}
cellchat <- computeCommunProbPathway(cellchat)
```

```{r}
cellchat <- aggregateNet(cellchat)
```

```{r}
df.pathways <- subsetCommunication(cellchat, slot.name = "netP", thresh = 0.05)
```

### Aggregated cell-cell communication network

```{r, fig.width=9, fig.height=3}
groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1,2), mar = c(0, 0, 2, 0), xpd=TRUE)
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")
```

### Cell-cell communication network split by sender

```{r, fig.width=9, fig.height=5}
mat <- cellchat@net$weight
par(mfrow = c(2,3), mar = c(0, 4, 0, 4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
```


### Interaction strength and number of interactions for each cell type pair

```{r}
netVisual_heatmap(cellchat, color.heatmap = "Reds", slot.name = "net", measure = "weight")
```

```{r}
netVisual_heatmap(cellchat, color.heatmap = "Reds", slot.name = "net", measure = "count")
```


### Dominant senders (sources) and receivers (targets)

```{r}
# Run the analysis without parallel processing
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP")

gg1 <- netAnalysis_signalingRole_scatter(cellchat)
gg1
```


### Signals contributing the most to outgoing or incoming signaling

```{r, fig.height=13, fig.width=11}
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", width = 10,
  height = 25)
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", width = 10,
  height = 25)
ht1 + ht2
```

```{r, fig.height=6, fig.width=8}
df <- rankNet(cellchat, mode = "single", stacked = T, return.data = T)$signaling.contribution %>%
  arrange(desc(contribution))
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", width = 7,
  height = 8, signaling = df[1:25,]$name)
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", width = 7,
  height = 8, signaling = df[1:25,]$name)
ht1 + ht2
```


### Top interactions between all cell types

```{r, fig.height=6, fig.width=6}
net <- df.net %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 1, legend.pos.y = 130, legend.pos.x = 5, slot.name = "net", net = net[1:25,])
```

```{r, fig.height=6, fig.width=6}
net <- df.pathways %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 0.65, legend.pos.y = 110, legend.pos.x = 5, slot.name = "netP", net = net[1:25,], small.gap = 2)
```


### Top interactions with endothelial cells as senders

```{r, fig.height=6, fig.width=6}
net <- df.net %>%
  filter(source == "Endothelial cells") %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 1, legend.pos.y = 130, legend.pos.x = 5, slot.name = "net", net = net[1:25,], sources.use = "Endothelial cells")
```

```{r, fig.height=8, fig.width=10}
net <- df.pathways %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, sources.use = "Endothelial cells", lab.cex = 1,legend.pos.y = 150, slot.name = "netP", net = net[1:100,])
```

```{r}
net <- df.net %>%
  filter(source == "Endothelial cells") %>%
  arrange(desc(prob))
df <- as.data.frame(net[1:25,]$interaction_name)
colnames(df) <- "interaction_name"
netVisual_bubble(cellchat, remove.isolate = FALSE, pairLR.use = df, sources.use = "Endothelial cells", angle.x = 45)
```


### Top interactions with endothelial cells as senders and cardiomyocytes as receivers

```{r, fig.height=6, fig.width=6}
net <- df.net %>%
  filter(source == "Endothelial cells") %>%
  filter(target %in% c("Ventricular Cardiomyocytes", "Atrial Cardiomyocytes")) %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 1, legend.pos.y = 135, legend.pos.x = 5, slot.name = "net", net = net[1:25,], sources.use = "Endothelial cells", targets.use = c("Ventricular Cardiomyocytes", "Atrial Cardiomyocytes"))
```

```{r}
df <- as.data.frame(net[1:25,]$interaction_name)
colnames(df) <- "interaction_name"
netVisual_bubble(cellchat, remove.isolate = FALSE, pairLR.use = df, sources.use = "Endothelial cells", angle.x = 45, targets.use = c("Ventricular Cardiomyocytes", "Atrial Cardiomyocytes"))
```


## Subset only Non-protein Signaling

```{r}
cellchat <- createCellChat(object = iPS_integrated, group.by = "ident", assay = "RNA")
cellchat <- addMeta(cellchat, meta = meta)
cellchat <- setIdent(cellchat, ident.use = "labels") # set "labels" as default cell identity
levels(cellchat@idents) # show factor levels of the cell labels
groupSize <- as.numeric(table(cellchat@idents)) # number of cells in each cell group
groupSize
```

```{r}

CellChatDB <- CellChatDB.human 
cellchat@DB <- CellChatDB
CellChatDB.use <- subsetDB(CellChatDB, search = "Non-protein Signaling", key = "annotation") # use Secreted Signaling
cellchat@DB <- CellChatDB.use
```

```{r}
# subset the expression data of signaling genes for saving computation cost
cellchat <- subsetData(cellchat, ) # This step is necessary even if using the whole database
#future::plan("multisession", workers = 4) # do parallel
cellchat <- identifyOverExpressedGenes(cellchat)
cellchat <- identifyOverExpressedInteractions(cellchat)
#> The number of highly variable ligand-receptor pairs used for signaling inference is 692
```

```{r}
cellchat <- computeCommunProb(cellchat, type = "triMean")
cellchat <- filterCommunication(cellchat, min.cells = 10)
```

```{r}
df.net <- subsetCommunication(cellchat, slot.name = "net")
```


```{r}
cellchat <- computeCommunProbPathway(cellchat)
```

```{r}
cellchat <- aggregateNet(cellchat)
```

```{r}
df.pathways <- subsetCommunication(cellchat, slot.name = "netP", thresh = 0.05)
```

### Aggregated cell-cell communication network

```{r, fig.width=9, fig.height=3}
groupSize <- as.numeric(table(cellchat@idents))
par(mfrow = c(1,2), mar = c(0, 0, 2, 0), xpd=TRUE)
netVisual_circle(cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")
```

### Cell-cell communication network split by sender

```{r, fig.width=9, fig.height=5}
mat <- cellchat@net$weight
par(mfrow = c(2,3), mar = c(0, 4, 0, 4), xpd=TRUE)
for (i in 1:nrow(mat)) {
  mat2 <- matrix(0, nrow = nrow(mat), ncol = ncol(mat), dimnames = dimnames(mat))
  mat2[i, ] <- mat[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat), title.name = rownames(mat)[i])
}
```


### Interaction strength and number of interactions for each cell type pair

```{r}
netVisual_heatmap(cellchat, color.heatmap = "Reds", slot.name = "net", measure = "weight")
```

```{r}
netVisual_heatmap(cellchat, color.heatmap = "Reds", slot.name = "net", measure = "count")
```


### Dominant senders (sources) and receivers (targets)

```{r}
# Run the analysis without parallel processing
cellchat <- netAnalysis_computeCentrality(cellchat, slot.name = "netP")

gg1 <- netAnalysis_signalingRole_scatter(cellchat)
gg1
```


### Signals contributing the most to outgoing or incoming signaling

```{r, fig.height=13, fig.width=11}
# Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", width = 10,
  height = 25)
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", width = 10,
  height = 25)
ht1 + ht2
```

```{r, fig.height=6, fig.width=8}
df <- rankNet(cellchat, mode = "single", stacked = T, return.data = T)$signaling.contribution %>%
  arrange(desc(contribution))
ht1 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "outgoing", width = 7,
  height = 8, signaling = df[1:25,]$name)
ht2 <- netAnalysis_signalingRole_heatmap(cellchat, pattern = "incoming", width = 7,
  height = 8, signaling = df[1:25,]$name)
ht1 + ht2
```


### Top interactions between all cell types

```{r, fig.height=6, fig.width=6}
net <- df.net %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 1, legend.pos.y = 130, legend.pos.x = 5, slot.name = "net", net = net[1:25,])
```

```{r, fig.height=6, fig.width=6}
net <- df.pathways %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 0.65, legend.pos.y = 110, legend.pos.x = 5, slot.name = "netP", net = net[1:25,], small.gap = 2)
```


### Top interactions with endothelial cells as senders

```{r, fig.height=6, fig.width=6}
net <- df.net %>%
  filter(source == "Endothelial cells") %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 1, legend.pos.y = 130, legend.pos.x = 5, slot.name = "net", net = net[1:25,], sources.use = "Endothelial cells")
```

```{r, fig.height=8, fig.width=10}
net <- df.pathways %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, sources.use = "Endothelial cells", lab.cex = 1,legend.pos.y = 150, slot.name = "netP", net = net[1:100,])
```

```{r}
net <- df.net %>%
  filter(source == "Endothelial cells") %>%
  arrange(desc(prob))
df <- as.data.frame(net[1:25,]$interaction_name)
colnames(df) <- "interaction_name"
netVisual_bubble(cellchat, remove.isolate = FALSE, pairLR.use = df, sources.use = "Endothelial cells", angle.x = 45)
```


### Top interactions with endothelial cells as senders and cardiomyocytes as receivers

```{r, fig.height=6, fig.width=6}
net <- df.net %>%
  filter(source == "Endothelial cells") %>%
  filter(target %in% c("Ventricular Cardiomyocytes", "Atrial Cardiomyocytes")) %>%
  arrange(desc(prob))
netVisual_chord_gene(cellchat, lab.cex = 1, legend.pos.y = 135, legend.pos.x = 5, slot.name = "net", net = net[1:25,], sources.use = "Endothelial cells", targets.use = c("Ventricular Cardiomyocytes", "Atrial Cardiomyocytes"))
```

```{r}
df <- as.data.frame(net[1:25,]$interaction_name)
colnames(df) <- "interaction_name"
netVisual_bubble(cellchat, remove.isolate = FALSE, pairLR.use = df, sources.use = "Endothelial cells", angle.x = 45, targets.use = c("Ventricular Cardiomyocytes", "Atrial Cardiomyocytes"))
```
