---
title: "QC_integration_annotation"
author: "GinoBonazza (ginoandrea.bonazza@usz.ch)"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r knitr config, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(warning = FALSE)

knitr::opts_chunk$set(message = FALSE)

knitr::opts_chunk$set(cache = FALSE)

knitr::opts_chunk$set(dpi = 600, fig.align = "center")
```

## Setup

```{r setup, class.source = "fold-hide"}
# Get current file name to make folder
current_file <- "QC_integration_annotation"

# Load libraries
library(here)
library(readr)
library(readxl)
library(xlsx)
library(Seurat)
library(DropletUtils)
library(Matrix)
library(scDblFinder)
library(scCustomize)
library(dplyr)
library(ggplot2)
library(magrittr)
library(tidyverse)
library(reshape2)
library(S4Vectors)
library(SingleCellExperiment)
library(pheatmap)
library(png)
library(gridExtra)
library(knitr)
library(scales)
library(RColorBrewer)
library(Matrix.utils)
library(tibble)
library(ggplot2)
library(scater)
library(patchwork)
library(statmod)
library(ArchR)
library(clustree)

#Output paths
output_dir_data <- here::here("output", current_file)
if (!dir.exists(output_dir_data)) dir.create(output_dir_data)

if (!dir.exists(here::here("docs", "figure"))) dir.create(here::here("docs", "figure"))

output_dir_figs <- here::here("docs", "figure", paste0(current_file, ".Rmd"))
if (!dir.exists(output_dir_figs)) dir.create(output_dir_figs)
```


## Quality Control

Load and merge cellranger output

```{r echo = TRUE, results= "hide"}
JD6_Ctr <- Read10X_h5(filename = here::here("data", "Cellranger_output", "JD6_Ctr_filtered_feature_bc_matrix.h5"))
JD6_Ctr <- CreateSeuratObject(JD6_Ctr, project = "JD6_Ctr")
JD6_DMD <- Read10X_h5(filename = here::here("data", "Cellranger_output", "JD6_DMD_filtered_feature_bc_matrix.h5"))
JD6_DMD <- CreateSeuratObject(JD6_DMD, project = "JD6_DMD")
DMD1_Ctr <- Read10X_h5(filename = here::here("data", "Cellranger_output", "DMD1_Ctr_filtered_feature_bc_matrix.h5"))
DMD1_Ctr <- CreateSeuratObject(DMD1_Ctr, project = "DMD1_Ctr")
DMD1 <- Read10X_h5(filename = here::here("data", "Cellranger_output", "DMD1_filtered_feature_bc_matrix.h5"))
DMD1 <- CreateSeuratObject(DMD1, project = "DMD1")
JDR_Ctr <- Read10X_h5(filename = here::here("data", "Cellranger_output", "JDR_Ctr_filtered_feature_bc_matrix.h5"))
JDR_Ctr <- CreateSeuratObject(JDR_Ctr, project = "JDR_Ctr")
JDR_DMD <- Read10X_h5(filename = here::here("data", "Cellranger_output", "JDR_DMD_filtered_feature_bc_matrix.h5"))
JDR_DMD <- CreateSeuratObject(JDR_DMD, project = "JDR_DMD")

iPS_preQC <- merge(JD6_Ctr, y = c(JD6_DMD, DMD1_Ctr, DMD1, JDR_Ctr, JDR_DMD), 
                   add.cell.ids = c("JD6_Ctr", "JD6_DMD", "DMD1_Ctr", "DMD1", "JDR_Ctr", "JDR_DMD"))
rm(JD6_Ctr, JD6_DMD, DMD1_Ctr, DMD1, JDR_Ctr, JDR_DMD)
```

Add metadata: percentage of mitochondrial and ribosomal genes, cell line, condition

```{r}
iPS_preQC[["percent.mt"]] <- PercentageFeatureSet(iPS_preQC, pattern = "^MT-")
iPS_preQC[["percent.rp"]] <- PercentageFeatureSet(iPS_preQC, pattern = "^RP[SL]")
iPS_preQC$Sample <- iPS_preQC$orig.ident
Organoids_metadata <- read_excel(here::here("data", "Organoids_metadata.xlsx"))
add_metadata <- left_join(iPS_preQC[["Sample"]], Organoids_metadata)
row.names(add_metadata) <- row.names(iPS_preQC[[]])
iPS_preQC <- AddMetaData(iPS_preQC, metadata = add_metadata)
rm(add_metadata, Organoids_metadata)
metadata <- unique(subset(iPS_preQC@meta.data, select = c(orig.ident, Sample:length(iPS_preQC@meta.data))))
rownames(metadata) <- metadata$orig.ident
metadata
```

Check quality control parameters

```{r QC_pre_VlnPlots, fig.height=6, fig.width=8}
p1 <- VlnPlot(iPS_preQC, features = "nCount_RNA", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

p2 <- VlnPlot(iPS_preQC, features = "nFeature_RNA", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

p3 <- VlnPlot(iPS_preQC, features = "percent.mt", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

p4 <- VlnPlot(iPS_preQC, features = "percent.rp", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

QC_pre_VlnPlots <- p1 + p2 + p3 + p4 + plot_layout(ncol = 2)

QC_pre_VlnPlots
```

```{r QC_pre_VlnPlots_zoom, fig.height=10, fig.width=8}
p1 <- VlnPlot(iPS_preQC, features = "nCount_RNA", group.by = "Sample", pt.size = 0, y.max = 10000) + theme(axis.title.x = element_blank()) + NoLegend()

p2 <- VlnPlot(iPS_preQC, features = "nFeature_RNA", group.by = "Sample", pt.size = 0, y.max = 6000) + theme(axis.title.x = element_blank()) + NoLegend()

p3 <- VlnPlot(iPS_preQC, features = "percent.mt", group.by = "Sample", pt.size = 0, y.max = 25) + theme(axis.title.x = element_blank()) + NoLegend()

p4 <- VlnPlot(iPS_preQC, features = "percent.rp", group.by = "Sample", pt.size = 0, y.max = 20) + theme(axis.title.x = element_blank()) + NoLegend()

QC_pre_VlnPlots_zoom <- p1 + p2 + p3 + p4 + plot_layout(ncol = 2)

QC_pre_VlnPlots_zoom
```

Doublets detection

```{r}
iPS_preQC_sce <- as.SingleCellExperiment(iPS_preQC)
iPS_preQC_sce <- scDblFinder(iPS_preQC_sce, samples="Sample", clusters = TRUE)
table(iPS_preQC_sce@colData$scDblFinder.class)
iPS_preQC <- as.Seurat(iPS_preQC_sce, counts = "counts", data = "logcounts")
rm(iPS_preQC_sce)
iPS_preQC@meta.data[iPS_preQC@meta.data$scDblFinder.class %in% "singlet", "scDblFinder.n"] = paste0("Singlets (n=", table(iPS_preQC$scDblFinder.class)[1], ")")
iPS_preQC@meta.data[iPS_preQC@meta.data$scDblFinder.class %in% "doublet", "scDblFinder.n"] = paste0("Doublets (n=", table(iPS_preQC$scDblFinder.class)[2], ")")
iPS_preQC$scDblFinder.n <- factor(x = iPS_preQC$scDblFinder.n, levels = c(rownames(table(iPS_preQC$scDblFinder.n))[2], rownames(table(iPS_preQC$scDblFinder.n))[1])) 
```

```{r QC_pre_Doublets, fig.height=3, fig.width=8}
p1 <- VlnPlot(iPS_preQC, features = c("nCount_RNA"), split.by = "scDblFinder.n", group.by = "Sample", pt.size = 0, y.max = 50000) + theme(axis.title.x = element_blank()) + NoLegend()

p2 <- VlnPlot(iPS_preQC, features = c("nFeature_RNA"), split.by = "scDblFinder.n", group.by = "Sample", pt.size = 0, y.max = 10000) + theme(axis.title.x = element_blank()) 

QC_pre_Doublets <- p1 + p2 + plot_layout(ncol = 2)

QC_pre_Doublets
```

```{r eval=FALSE}
saveRDS(iPS_preQC, 
        here::here(output_dir_data, "iPS_preQC.rds"))
```

Remove doublets

```{r iPS_preQC_no_doublets}
iPS_preQC_no_doublets <- subset(x = iPS_preQC, subset = scDblFinder.class == "singlet")
table(iPS_preQC_no_doublets@meta.data$scDblFinder.class)
rm(iPS_preQC)
```

Check quality control parameters after removing doublets

```{r QC_pre_no_doublets_VlnPlots, fig.height=6, fig.width=8}
p1 <- VlnPlot(iPS_preQC_no_doublets, features = "nCount_RNA", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

p2 <- VlnPlot(iPS_preQC_no_doublets, features = "nFeature_RNA", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

p3 <- VlnPlot(iPS_preQC_no_doublets, features = "percent.mt", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

p4 <- VlnPlot(iPS_preQC_no_doublets, features = "percent.rp", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

QC_pre_no_doublets_VlnPlots <- p1 + p2 + p3 + p4 + plot_layout(ncol = 2)

QC_pre_no_doublets_VlnPlots
```

```{r QC_pre_no_doublets_VlnPlots_zoom, fig.height=10, fig.width=8}
p1 <- VlnPlot(iPS_preQC_no_doublets, features = "nCount_RNA", group.by = "Sample", pt.size = 0, y.max = 10000) + theme(axis.title.x = element_blank()) + NoLegend()

p2 <- VlnPlot(iPS_preQC_no_doublets, features = "nFeature_RNA", group.by = "Sample", pt.size = 0, y.max = 6000) + theme(axis.title.x = element_blank()) + NoLegend()

p3 <- VlnPlot(iPS_preQC_no_doublets, features = "percent.mt", group.by = "Sample", pt.size = 0, y.max = 25) + theme(axis.title.x = element_blank()) + NoLegend()

p4 <- VlnPlot(iPS_preQC_no_doublets, features = "percent.rp", group.by = "Sample", pt.size = 0, y.max = 20) + theme(axis.title.x = element_blank()) + NoLegend()

QC_pre_no_doublets_VlnPlots_zoom <- p1 + p2 + p3 + p4 + plot_layout(ncol = 2)

QC_pre_no_doublets_VlnPlots_zoom
```

Filter based on number of counts, features and percentage of mitochondrial genes

```{r}
iPS <- subset(iPS_preQC_no_doublets, subset = 
                nFeature_RNA > 600 &
                percent.mt < 20 &
                nCount_RNA > 1000)

table(iPS$Sample)
```

Check quality control parameters

```{r QC_post_VlnPlots, fig.height=6, fig.width=8}
p1 <- VlnPlot(iPS, features = "nCount_RNA", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

p2 <- VlnPlot(iPS, features = "nFeature_RNA", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

p3 <- VlnPlot(iPS, features = "percent.mt", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

p4 <- VlnPlot(iPS, features = "percent.rp", group.by = "Sample", pt.size = 0) + theme(axis.title.x = element_blank()) + NoLegend()

QC_post_VlnPlots <- p1 + p2 + p3 + p4 + plot_layout(ncol = 2)

QC_post_VlnPlots
```

```{r QC_post_VlnPlots_zoom, fig.height=10, fig.width=8}
p1 <- VlnPlot(iPS, features = "nCount_RNA", group.by = "Sample", pt.size = 0, y.max = 10000) + theme(axis.title.x = element_blank()) + NoLegend()

p2 <- VlnPlot(iPS, features = "nFeature_RNA", group.by = "Sample", pt.size = 0, y.max = 6000) + theme(axis.title.x = element_blank()) + NoLegend()

p3 <- VlnPlot(iPS, features = "percent.mt", group.by = "Sample", pt.size = 0, y.max = 25) + theme(axis.title.x = element_blank()) + NoLegend()

p4 <- VlnPlot(iPS, features = "percent.rp", group.by = "Sample", pt.size = 0, y.max = 20) + theme(axis.title.x = element_blank()) + NoLegend()

QC_post_VlnPlots_zoom <- p1 + p2 + p3 + p4 + plot_layout(ncol = 2)

QC_post_VlnPlots_zoom
```


## Integration and clustering

Normalization and scaling using SCTransform.

```{r eval=FALSE}
DefaultAssay(iPS) <- "RNA"
iPS.list <- SplitObject(iPS, split.by = "Sample")
for (i in 1:length(iPS.list)) {
  iPS.list[[i]] <- SCTransform(iPS.list[[i]], vst.flavor = "v2", vars.to.regress = c("percent.mt", "percent.rp"))
}
```

PCA without integration

```{r, eval=FALSE}
features <- SelectIntegrationFeatures(object.list = iPS.list, nfeatures = 2000)
iPS_not_integrated <- Merge_Seurat_List(iPS.list)
VariableFeatures(iPS_not_integrated) <- features
iPS_not_integrated <- RunPCA(iPS_not_integrated)
```

```{r eval=FALSE}
saveRDS(iPS_not_integrated, 
        here::here(output_dir_data, "iPS_not_integrated.rds"))
```

```{r}
iPS_not_integrated <- readRDS(here::here(output_dir_data, "iPS_not_integrated.rds"))
```


```{r Elbow_not_integrated, fig.width=4, fig.height=4}
ElbowPlot(iPS_not_integrated, ndims = 50)
```

Clustering without integration

```{r eval=FALSE}
iPS_not_integrated <- RunUMAP(iPS_not_integrated, dims = 1:40)
iPS_not_integrated <- FindNeighbors(iPS_not_integrated, dims = 1:40)
iPS_not_integrated <- FindClusters(iPS_not_integrated, resolution = seq(0.1, 0.8, by=0.1))
```

```{r Clustree_not_integrated, fig.width=6, fig.height=9}
clustree::clustree(iPS_not_integrated@meta.data[,grep("SCT_snn_res", colnames(iPS_not_integrated@meta.data))],
                   prefix = "SCT_snn_res.")
```

```{r UMAPs_not_integrated, fig.width=12, fig.height=10}
DimPlot(iPS_not_integrated, reduction = "umap", shuffle = T,
        group.by = c("SCT_snn_res.0.3", "Sample", "Condition", "Cell_line"), ncol = 2)
```

```{r eval=FALSE}
saveRDS(iPS_not_integrated, 
        here::here(output_dir_data, "iPS_not_integrated.rds"))
```

Integrate the samples

```{r eval=FALSE}
iPS.list <- PrepSCTIntegration(object.list = iPS.list, anchor.features = features)
iPS.anchors <- FindIntegrationAnchors(object.list = iPS.list, normalization.method = "SCT",
                                      anchor.features = features)
iPS_integrated <- IntegrateData(anchorset = iPS.anchors, normalization.method = "SCT")
```

PCA after integration

```{r eval=FALSE}
iPS_integrated <- RunPCA(iPS_integrated)
```

```{r, eval=FALSE}
saveRDS(iPS_integrated, 
        here::here(output_dir_data, "iPS_integrated_low_quality.rds"))
```

```{r}
iPS_integrated <- readRDS(here::here(output_dir_data, "iPS_integrated_low_quality.rds"))
```

```{r Elbow_integrated_low_quality, fig.width=4, fig.height=4}
ElbowPlot(iPS_integrated, ndims = 50)
```

Clustering after integration

```{r eval=FALSE}
DefaultAssay(iPS_integrated) <- "integrated"
iPS_integrated <- RunUMAP(iPS_integrated, dims = 1:30, reduction = "pca")
iPS_integrated <- FindNeighbors(iPS_integrated, dims = 1:30, reduction = "pca")
iPS_integrated <- FindClusters(iPS_integrated, resolution = seq(0.1, 0.8, by=0.1))
```

```{r Clustree_integrated_low_quality, fig.width=6, fig.height=9}
clustree::clustree(iPS_integrated@meta.data[,grep("integrated_snn_res", colnames(iPS_integrated@meta.data))],
                   prefix = "integrated_snn_res.")
```

```{r UMAPs_integrated_low_quality, fig.width=12, fig.height=10}
DimPlot(iPS_integrated, reduction = "umap", shuffle = T,
        group.by = c("integrated_snn_res.0.2", "Sample", "Condition", "Cell_line"), ncol = 2)
```

```{r, eval=FALSE}
saveRDS(iPS_integrated, 
        here::here(output_dir_data, "iPS_integrated_low_quality.rds"))
```

Find the markers that characterize each cell population

```{r}
DefaultAssay(iPS_integrated) <- "SCT"
iPS_integrated <- PrepSCTFindMarkers(iPS_integrated)
Idents(iPS_integrated) <- "integrated_snn_res.0.2"
Markers <- FindAllMarkers(iPS_integrated, only.pos = TRUE, min.pct = 0.5, logfc.threshold = 0.5)
write.csv(Markers, here::here(output_dir_data, "iPS_Markers_all_low_quality.csv"))
Markers_top10 <- as.data.frame(Markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC))
write.csv(Markers_top10, here::here(output_dir_data, "iPS_Markers_top10_low_quality.csv"))
Markers_top3 <- as.data.frame(Markers %>% group_by(cluster) %>% top_n(n = 3, wt = avg_log2FC))
write.csv(Markers_top3, here::here(output_dir_data, "iPS_Markers_top3_low_quality.csv"))
```

```{r Heatmap_low_quality, fig.width=12, fig.height=10}
DefaultAssay(iPS_integrated) <- "RNA"
iPS_integrated <- NormalizeData(iPS_integrated)
iPS_integrated <- ScaleData(iPS_integrated)
mapal <- colorRampPalette(RColorBrewer::brewer.pal(9,"RdBu"))(256)
mapal <- rev(mapal[1:256])
Heatmap <- DoHeatmap(iPS_integrated, draw.line = F, features = Markers_top10$gene) +
  guides(colour=FALSE) +
  scale_fill_gradientn(colours = mapal) +
  theme(text = element_text(size = 15), axis.text.y = element_text(size = 8)) +
  theme(plot.margin = unit(c(0.1, 0, 0, 0), 
                           "inches"))
Heatmap
```

Check QC parameters in each cluster

```{r QC_VlnPlots_low_quality, fig.width=8, fig.height=6}
VlnPlot(iPS_integrated, features = c("nCount_RNA", "nFeature_RNA", "percent.mt", "percent.rp"), ncol = 2) + theme(axis.title.x = element_blank())
```

Remove low quality cells (clusters with low number of features)
```{r}
iPS_integrated <- subset(iPS_integrated, subset = integrated_snn_res.0.2 != "3" & integrated_snn_res.0.2 != "5")
```

Recluster

```{r eval=FALSE}
DefaultAssay(iPS_integrated) <- "integrated"
iPS_integrated <- RunPCA(iPS_integrated)
```

```{r eval=FALSE}
saveRDS(iPS_integrated, 
        here::here(output_dir_data, "iPS_integrated.rds"))
```

```{r}
iPS_integrated <- readRDS(here::here(output_dir_data, "iPS_integrated.rds"))
```

```{r Elbow_integrated, fig.width=4, fig.height=4}
ElbowPlot(iPS_integrated, ndims = 50)
```

```{r eval=FALSE}
DefaultAssay(iPS_integrated) <- "integrated"
iPS_integrated <- RunUMAP(iPS_integrated, dims = 1:30, reduction = "pca")
iPS_integrated <- FindNeighbors(iPS_integrated, dims = 1:30, reduction = "pca")
iPS_integrated <- FindClusters(iPS_integrated, resolution = seq(0.1, 0.8, by=0.1))
```

```{r Clustree_integrated, fig.width=6, fig.height=9}
clustree::clustree(iPS_integrated@meta.data[,grep("integrated_snn_res", colnames(iPS_integrated@meta.data))],
                   prefix = "integrated_snn_res.")
```

```{r UMAPs_integrated, fig.width=12, fig.height=10}
DimPlot(iPS_integrated, reduction = "umap", shuffle = T,
        group.by = c("integrated_snn_res.0.2", "Sample", "Condition", "Cell_line"), ncol = 2)
```

Find the markers that characterize each cell population

```{r}
DefaultAssay(iPS_integrated) <- "SCT"
iPS_integrated <- PrepSCTFindMarkers(iPS_integrated)
Idents(iPS_integrated) <- "integrated_snn_res.0.2"
Markers <- FindAllMarkers(iPS_integrated, only.pos = TRUE, min.pct = 0.5, logfc.threshold = 0.5, recorrect_umi = FALSE)
write.csv(Markers, here::here(output_dir_data, "iPS_Markers_all.csv"))
Markers_top10 <- as.data.frame(Markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC))
write.csv(Markers_top10, here::here(output_dir_data, "iPS_Markers_top10.csv"))
Markers_top3 <- as.data.frame(Markers %>% group_by(cluster) %>% top_n(n = 3, wt = avg_log2FC))
write.csv(Markers_top3, here::here(output_dir_data, "iPS_Markers_top3.csv"))
```

Check cell cycle genes
```{r Cell_cycle_UMAP, fig.width=5, fig.height=4}
DefaultAssay(iPS_integrated) <- "RNA"
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
iPS_integrated <- CellCycleScoring(iPS_integrated, s.features = s.genes, g2m.features = g2m.genes)
DimPlot(iPS_integrated, reduction = "umap", shuffle = T,
        group.by = "Phase")
```


## Cell type annotation

Annotate the clusters based on the characteristic markers

```{r}
iPS_integrated@meta.data[iPS_integrated@meta.data$integrated_snn_res.0.2 %in% c("0"), "cell_type"] = "Atrial Cardiomyocytes"
iPS_integrated@meta.data[iPS_integrated@meta.data$integrated_snn_res.0.2 %in% c("1"), "cell_type"] = "Ventricular Cardiomyocytes"
iPS_integrated@meta.data[iPS_integrated@meta.data$integrated_snn_res.0.2 %in% c("2"), "cell_type"] = "CDKN1A+ Cardiomyocytes"
iPS_integrated@meta.data[iPS_integrated@meta.data$integrated_snn_res.0.2 %in% c("3"), "cell_type"] = "Fibroblasts"
iPS_integrated@meta.data[iPS_integrated@meta.data$integrated_snn_res.0.2 %in% c("4"), "cell_type"] = "Progenitor cells"
iPS_integrated@meta.data[iPS_integrated@meta.data$integrated_snn_res.0.2 %in% c("5"), "cell_type"] = "Endothelial cells"
iPS_integrated$cell_type <- factor(iPS_integrated$cell_type, levels = c("Atrial Cardiomyocytes", 
                                                                        "Ventricular Cardiomyocytes",
                                                                        "CDKN1A+ Cardiomyocytes",
                                                                        "Fibroblasts",
                                                                        "Progenitor cells",
                                                                        "Endothelial cells"))
Idents(iPS_integrated) <- iPS_integrated$cell_type
```

```{r Heatmap, fig.width=12, fig.height=10}
DefaultAssay(iPS_integrated) <- "RNA"
cluster_palette <- c("#24B700","#F8766D", "#D575FE", "#00ACFC", "#EA8331", "#FF65AC")
Heatmap <- DoHeatmap(iPS_integrated, draw.line = F, features = Markers_top10$gene, group.colors = cluster_palette) +
  guides(colour=FALSE) +
  scale_fill_gradientn(colours = mapal) +
  theme(text = element_text(size = 15), axis.text.y = element_text(size = 8)) +
  theme(plot.margin = unit(c(0, 0, 0, 0), 
                           "inches"))
Heatmap
```

```{r UMAP_cell_type, fig.width=6.5, fig.height=6}
p <- DimPlot(iPS_integrated, group.by = "cell_type", reduction = "umap", label = F, cols = cluster_palette) + 
  NoLegend() + 
  theme(axis.text=element_text(size=14, face = "bold"), axis.title = element_text(size = 18, face = "bold")) + 
  theme(plot.title = element_blank())
LabelClusters(p, id = "cell_type", fontface = "bold", size = 5.5, repel = T)
```

```{r DotPlot_markers, fig.width=6, fig.height=7}
DotPlot(iPS_integrated, assay = "SCT", features = rev(Markers_top3$gene), dot.scale = 5, cluster.idents = FALSE) +
  RotatedAxis() +
  coord_flip() +
  theme(axis.title = element_blank(), axis.text.x = element_text(size = 14), axis.text.y = element_text(size = 13), legend.text = element_text(size = 9), legend.title = element_text(size = 11), plot.margin = unit(c(0, 0, 0, 0.1), 
                           "inches"))
```

```{r FeatPlot_markers, fig.width=13.5, fig.height=8}
FeaturePlot(iPS_integrated, features = c("MYH6", "MYH7", "CDKN1A", "FN1", "TOP2A", "PECAM1"), ncol = 3)
```

```{r, eval = FALSE}
saveRDS(iPS_integrated, 
        here::here(output_dir_data, "iPS_integrated.rds"))
```






